Abstract Plots

### 2022-02-01 Abstract.Rmd
Evaluate max/min sensitivity, specificity, and AUC among diluted samples at which classifier fails for a fixed optimal sensitivity/specificity/AUC threshold and for all COSMIC signatures. Each evaluation is plotted and includes a linear model with distributions of optimal sensitivity/specificity/AUC.

---
title: "Abstract Plots"
author: "Audrey Mitchell"
date: "7/21/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

@@

```{r}
vector_to_matrix <- function(mutations.vector){
  tab <- table(mutations.vector)/sum(table(mutations.vector))
  #transform table to a data frame and assign column names to mutations and frequencies
  mutations.df <- data.frame(tab)
  colnames(mutations.df) <- c("mutations", "frequencies")

@@

```{r}
#load FFPE signature as a matrix
ffpe.signature <- as.matrix(load_old_mutational_matrix("supplied_results/ffpe.signature.txt"))

@@

```{r}
mmr.sample <- create_signature_sample_vector(signature.6.mat, 200)
tobacco.sample <- create_signature_sample_vector(signature.4.mat, 200)
ffpe.sample <- create_signature_sample_vector(ffpe.signature, 100)

@@

ggsave("mutprofile.png", plot = mut.profile , width = 175, height = 108, units = "mm")
```

```{r}
sig.mat <- cbind(signature.4.mat, signature.6.mat, ffpe.signature)
colnames(sig.mat) <- c("Signature 4", "Signature 6", "FFPE")
fts.res <- fit_to_signatures(ex.ffpe.profile.mat, sig.mat)

@@

ggsave("contributions.png", plot = contribution.plot , width = 175, height = 54, units = "mm")
```

```{r}
signatures.plot <- plot_96_profile(sig.mat)+
  labs(title = "Mutational Signatures", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))

@@

ggsave("signatures.png", plot = signatures.plot , width = 175, height = 108, units = "mm")
```

# Diagram of Simulating Data
```{r}
signature4.plot <- plot_96_profile(signature.4.mat)+
  labs(title = "Signature 4", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))

ggsave("signature4.png", plot = signature4.plot , width = 175, height = 81, units = "mm")
```

```{r}
ffpe.plot <- plot_96_profile(ffpe.signature)+
  labs(title = "FFPE Signature", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))

ggsave("ffpe.png", plot = ffpe.plot , width = 175, height = 81, units = "mm")
```

```{r}
sample.4 <- create_signature_sample_vector(signature.4.mat, 500)
sample.ffpe <- create_signature_sample_vector(ffpe.signature, 500)

@@

ggsave("sampleprofile.png", plot = sampleprof.plot , width = 175, height = 108, units = "mm")
```

@@

```{r}
set.seed(500)
#Create Signature 4 Sample
sig4.500 <- create_signature_sample_vector(signature.4.mat, 500)
sig4.profile <- vector_to_matrix(sig4.500)

@@

```{r}
#Add FFPE Noise
addffpe.500 <- create_signature_sample_vector(ffpe.signature, 500)

@@

```{r}
#Classify and Filter
sig4.ffpe.class <- get_classification_df(list(sig4.500, addffpe.500), c("SBS4", "FFPE"), list(signature.4.mat, ffpe.signature))

@@

```{r}
#Calculate Sensitivity, Specificity, and Misclassification rate

@@

```{r}
#Arrange 3 plots
sig4.reconstruction <- grid.arrange(sig4.profile.plot, sig4.ffpe.plot, sig4.recon.plot)

ggsave("sig4reconstruction.png", plot = sig4.reconstruction , width = 175, height = 243, units = "mm")
```

```{r}
#Create Signature 6 Sample
sig6.500 <- create_signature_sample_vector(signature.6.mat, 500)
sig6.profile <- vector_to_matrix(sig6.500)

@@

```{r}
#Add FFPE Noise
sig6.ffpe <- vector_to_matrix(c(sig6.500, addffpe.500))

@@

```{r}
#Classify and Filter
sig6.ffpe.class <- get_classification_df(list(sig6.500, addffpe.500), c("SBS6", "FFPE"), list(signature.6.mat, ffpe.signature))

@@

```{r}
#Calculate Sensitivity, specificity, and misclassification rate

@@

```{r}
#Arrange 3 plots
sig6.reconstruction <- grid.arrange(sig6.profile.plot, sig6.ffpe.plot, sig6.recon.plot)

ggsave("sig6reconstruction.png", plot = sig6.reconstruction , width = 175, height = 243, units = "mm")
```

@@

```{r}
sample.sig6 <- create_signature_sample_vector(signature.6.mat, 500)
sig6.truth <- rep("MMR", 500)
sample.ffpe <- create_signature_sample_vector(ffpe.signature, 500)
ffpe.truth <- rep("FFPE", 500)
MMR.ROC <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"),
                              list(signature.matrix.list[[6]], ffpe.signature))

@@

```{r}
MMR.ROC <- MMR.ROC %>%
  #create binary labels
  mutate(truthMMR = ifelse(truth == "MMR",1,0))

@@

# Scatterplot of Average AUC vs. Cosine Similarity to FFPE
```{r}
set.seed(500)
#outer for loop: loop through each COSMIC v3 signature
average.auc.values <- c()
for (x in 1:length(signature.names)){
  signature.auc.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],500)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
    classification.df.1 <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df.1)
    sig.name <- class.df.colnames[3]
    classification.df.2 <- classification.df.1 %>%
      mutate(signature.indicator = ifelse(truth == sig.name, 1, 0))
    predictions <- classification.df.2[3]
    labels <- classification.df.2$signature.indicator
    pred <- prediction(predictions, labels)
    auc.perf <- performance(pred, measure = "auc")
    signature.auc.vec[y] <- auc.perf@y.values[[1]][1]
  }
  average.signature.auc <- mean(signature.auc.vec)
  average.auc.values[x] <- average.signature.auc
}
```

```{r}
cos.sim.auc.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.auc.values)
colnames(cos.sim.auc.data) <- c("signature", "cosine_similarity", "average_auc")
```

```{r}

ggplot(cos.sim.auc.data, aes(x=average_auc)) +
  geom_histogram(color="black", fill="lightblue")+
  labs( x = "Average AUC", y = "Count")
ggsave("supp.fig.5.png", dpi = 300, width = 175, height = 108, units = "mm")

@@

```{r}
ggplot(cos.sim.auc.data, aes(x = cosine_similarity, y = average_auc))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs( x = "Cosine similarity between baseline COSMIC and FFPE signatures", y = "Average AUC")+
  geom_point(data = cos.sim.auc.outliers, color = "red")+
  geom_point(data = cos.sim.auc.outliers2, color = "purple")+
  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+
  geom_hline(yintercept=0.9, linetype="dashed", color = "purple")+
  geom_label_repel(data = cos.sim.auc.outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
  geom_label_repel(data = cos.sim.auc.outliers2, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
  geom_label_repel(data = cos.sim.sbs4, aes(label = signature), max.overlaps = 10, label.size =   0.1)+

ggsave("Fig2b.png", dpi = 300, width = 175, height = 108, units = "mm")

@@

```{r}
set.seed(500)
#outer for loop: loop through each COSMIC v3 signature
average.sensitivity.values <- c()
average.specificity.values <- c()
for (x in 1:length(signature.names)){
  signature.sensitivity.vec <- c()
  signature.specificity.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
    classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df)
    sig.name <- class.df.colnames[3]
    #Count number of true positives (for sensititivity)
    sensitivity.df <- classification.df %>%
      filter(truth == sig.name & classify == sig.name)
    #Calculate sensitivity
    sensitivity.val = nrow(sensitivity.df)/100
    #Add to the vector of 10 sensitivity values
    signature.sensitivity.vec[y] = sensitivity.val
    #Count number of true negatives (for specificity)
    specificity.df <- classification.df %>%
      filter(truth == "FFPE" & classify == "FFPE")
    #Calculate specificity
    specificity.val = nrow(specificity.df)/100
    #Add to the vector of 10 specificity values
    signature.specificity.vec[y] = specificity.val
  }
  average.signature.sensitivity <- mean(signature.sensitivity.vec)
  average.sensitivity.values[x] <- average.signature.sensitivity
  average.signature.specificity <- mean(signature.specificity.vec)
  average.specificity.values[x] <- average.signature.specificity
}
```

```{r}
sens.spec.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.sensitivity.values, average.specificity.values)

@@

ggplot(sens.spec.data, aes(x=sensitivity)) +
  geom_histogram(color="black", fill="lightblue")+
  labs( x = "Average sensitivity", y = "Count")
ggsave("supp.fig.3.png", dpi = 300, width = 175, height = 108, units = "mm")

@@

ggplot(sens.spec.data, aes(x = cosine_similarity, y = sensitivity))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs( x = "Cosine similarity between baseline COSMIC and FFPE signatures", y = "Average sensitivity")+
  geom_point(data = outliers, color = "red")+
  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+
  geom_label_repel(data = outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)+

ggsave("supp.fig6.png", dpi = 300, width = 175, height = 108, units = "mm")

@@

ggplot(sens.spec.data, aes(x=specificity)) +
  geom_histogram(color="black", fill="lightblue")+
  labs( x = "Average specificity", y = "Count")
ggsave("supp.fig.4.png", dpi = 300, width = 175, height = 108, units = "mm")

@@

ggplot(sens.spec.data, aes(x = cosine_similarity, y = specificity))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs( x = "Cosine similarity between baseline COSMIC and FFPE signatures", y = "Average specificity")+
  geom_point(data = outliers, color = "red")+
  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+
  geom_label_repel(data = outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)+

ggsave("supp.fig7.png", dpi = 300, width = 175, height = 108, units = "mm")

@@

```{r}
num.mut = 1000
step = seq(50, 950, by = 50)
```

@@

```{r}
sens.MMR.vec = evaluate_perf(100, length(step), test.MMR, "sens")
```

```{r}
sens.tob.vec = evaluate_perf(100, length(step), test.tob, "sens")
```

@@

MMR.tob.sens <- ggplot(data = sens.df, mapping = aes(x = dilution, 
                                     y = sensitivity, 
                                     color = signature.sens)) +
  geom_point() +
  geom_smooth(se=FALSE) +
  labs(x = "Percent of FFPE signature",
       y = "Sensitivity",
       color = "Baseline signature") +
  coord_cartesian(xlim = c(0,100), ylim = c(0,1))+
  scale_y_continuous(breaks=seq(from=0, to=1, by=.2))+
  geom_hline(yintercept=0.8, linetype="dashed", color = "red")

ggsave("fig3a.png", plot = MMR.tob.sens, dpi = 300, width = 175, height = 108, units = "mm")

@@

@@

```{r}
specificity = append(spec.MMR.vec, spec.tob.vec)
signature.spec = append(rep("SBS6", length(sens.MMR.vec)),
                        rep("SBS4", length(spec.tob.vec)))
dilution = append(step/10, step/10)
spec.df <- data.frame(dilution, specificity, signature.spec)

MMR.tob.spec <- ggplot(data = spec.df, mapping = aes(x = dilution, 
                                     y = specificity, 
                                     color = signature.sens)) +
  geom_point() +
  geom_smooth() +
  labs(x = "Percent of FFPE signature",
       y = "Specificity",
       color = "Baseline signature") +
  coord_cartesian(xlim = c(0,100), ylim = c(0,1))+
    scale_y_continuous(breaks=seq(from=0, to=1, by=.2))+
  geom_hline(yintercept=0.8, linetype="dashed", color = "red")

ggsave("fig3b.png", plot = MMR.tob.spec, dpi = 300, width = 175, height = 108, units = "mm")

@@

@@

```{r}
dilution.sig = signature.6.mat
dilution.sig.name = "SBS6"
total.mut = 2000
dilute = c(0.1,0.3,0.5,0.7,0.9)*total.mut
dilution.list <- vector(mode = "list", length = 5)
i = 1

@@

ggsave("sig6reconstruction_1.png", plot = dilution.list[[1]], dpi = 300, width = 175, height = 243, units = "mm")
ggsave("sig6reconstruction_2.png", plot = dilution.list[[2]], dpi = 300, width = 175, height = 243, units = "mm")
ggsave("sig6reconstruction_3.png", plot = dilution.list[[3]], dpi = 300, width = 175, height = 243, units = "mm")
ggsave("sig6reconstruction_4.png", plot = dilution.list[[4]], dpi = 300, width = 175, height = 243, units = "mm")
ggsave("sig6reconstruction_5.png", plot = dilution.list[[5]], dpi = 300, width = 175, height = 243, units = "mm")
```

@@

```{r}
#columns in data frame
sens = vector(mode = "list", length = length(signature.names))
spec = vector(mode = "list", length = length(signature.names))
acc = vector(mode = "list", length = length(signature.names))
signature = vector(mode = "list", length = length(signature.names))

>>>>

---
title: "Abstract Plots"
author: "Audrey Mitchell"
date: "7/21/2021"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

@@

```{r, include = FALSE}
vector_to_matrix <- function(mutations.vector){
  tab <- table(mutations.vector)/sum(table(mutations.vector))
  #transform table to a data frame and assign column names to mutations and frequencies
  mutations.df <- data.frame(tab)
  colnames(mutations.df) <- c("mutations", "frequencies")

@@

```{r, include=FALSE}
#load FFPE signature as a matrix
ffpe.signature <- as.matrix(load_old_mutational_matrix("supplied_results/ffpe.signature.txt"))

@@

```{r, include=FALSE}
mmr.sample <- create_signature_sample_vector(signature.6.mat, 200)
tobacco.sample <- create_signature_sample_vector(signature.4.mat, 200)
ffpe.sample <- create_signature_sample_vector(ffpe.signature, 100)

@@

#ggsave("mutprofile.png", plot = mut.profile , width = 175, height = 108, units = "mm")
```

```{r, include=FALSE}
sig.mat <- cbind(signature.4.mat, signature.6.mat, ffpe.signature)
colnames(sig.mat) <- c("Signature 4", "Signature 6", "FFPE")
fts.res <- fit_to_signatures(ex.ffpe.profile.mat, sig.mat)

@@

#ggsave("contributions.png", plot = contribution.plot , width = 175, height = 54, units = "mm")
```

```{r, include=FALSE}
signatures.plot <- plot_96_profile(sig.mat)+
  labs(title = "Mutational Signatures", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))

@@

#ggsave("signatures.png", plot = signatures.plot , width = 175, height = 108, units = "mm")
```

# Diagram of Simulating Data
```{r, include=FALSE}
signature4.plot <- plot_96_profile(signature.4.mat)+
  labs(title = "Signature 4", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))

#ggsave("signature4.png", plot = signature4.plot , width = 175, height = 81, units = "mm")
```

```{r, include=FALSE}
ffpe.plot <- plot_96_profile(ffpe.signature)+
  labs(title = "FFPE Signature", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))

#ggsave("ffpe.png", plot = ffpe.plot , width = 175, height = 81, units = "mm")
```

```{r, include=FALSE}
sample.4 <- create_signature_sample_vector(signature.4.mat, 500)
sample.ffpe <- create_signature_sample_vector(ffpe.signature, 500)

@@

#ggsave("sampleprofile.png", plot = sampleprof.plot , width = 175, height = 108, units = "mm")
```

@@

```{r, include=FALSE}
set.seed(500)
#Create Signature 4 Sample
sig4.500 <- create_signature_sample_vector(signature.4.mat, 500)
sig4.profile <- vector_to_matrix(sig4.500)

@@

```{r, include=FALSE}
#Add FFPE Noise
addffpe.500 <- create_signature_sample_vector(ffpe.signature, 500)

@@

```{r, include=FALSE}
#Classify and Filter
sig4.ffpe.class <- get_classification_df(list(sig4.500, addffpe.500), c("SBS4", "FFPE"), list(signature.4.mat, ffpe.signature))

@@

```{r, include=FALSE}
#Calculate Sensitivity, Specificity, and Misclassification rate

@@

```{r, include=FALSE}
#Arrange 3 plots
sig4.reconstruction <- grid.arrange(sig4.profile.plot, sig4.ffpe.plot, sig4.recon.plot)

#ggsave("sig4reconstruction.png", plot = sig4.reconstruction , width = 175, height = 243, units = "mm")
```

```{r, include=FALSE}
#Create Signature 6 Sample
sig6.500 <- create_signature_sample_vector(signature.6.mat, 500)
sig6.profile <- vector_to_matrix(sig6.500)

@@

```{r, include=FALSE}
#Add FFPE Noise
sig6.ffpe <- vector_to_matrix(c(sig6.500, addffpe.500))

@@

```{r, include=FALSE}
#Classify and Filter
sig6.ffpe.class <- get_classification_df(list(sig6.500, addffpe.500), c("SBS6", "FFPE"), list(signature.6.mat, ffpe.signature))

@@

```{r, include=FALSE}
#Calculate Sensitivity, specificity, and misclassification rate

@@

```{r, include=FALSE}
#Arrange 3 plots
sig6.reconstruction <- grid.arrange(sig6.profile.plot, sig6.ffpe.plot, sig6.recon.plot)

#ggsave("sig6reconstruction.png", plot = sig6.reconstruction , width = 175, height = 243, units = "mm")
```

@@

```{r, include=FALSE}
sample.sig6 <- create_signature_sample_vector(signature.6.mat, 500)
sig6.truth <- rep("MMR", 500)
sample.ffpe <- create_signature_sample_vector(ffpe.signature, 500)
ffpe.truth <- rep("FFPE", 500)
MMR.ROC <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"),
                              list(signature.matrix.list[[6]], ffpe.signature))

@@

```{r, include=FALSE}
MMR.ROC <- MMR.ROC %>%
  #create binary labels
  mutate(truthMMR = ifelse(truth == "MMR",1,0))

@@

# Scatterplot of Average AUC vs. Cosine Similarity to FFPE
```{r, include=FALSE}
set.seed(500)
#outer for loop: loop through each COSMIC v3 signature
average.auc.values <- c()
for (x in 1:length(signature.names)){
  signature.auc.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],500)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
    classification.df.1 <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df.1)
    sig.name <- class.df.colnames[3]
    classification.df.2 <- classification.df.1 %>%
      mutate(signature.indicator = ifelse(truth == sig.name, 1, 0))
    predictions <- classification.df.2[3]
    labels <- classification.df.2$signature.indicator
    pred <- prediction(predictions, labels)
    auc.perf <- performance(pred, measure = "auc")
    signature.auc.vec[y] <- auc.perf@y.values[[1]][1]
  }
  average.signature.auc <- mean(signature.auc.vec)
  average.auc.values[x] <- average.signature.auc
}
```

```{r, include=FALSE}
cos.sim.auc.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.auc.values)
colnames(cos.sim.auc.data) <- c("signature", "cosine_similarity", "average_auc")
```

```{r, include=FALSE}

ggplot(cos.sim.auc.data, aes(x=average_auc)) +
  geom_histogram(color="black", fill="lightblue")+
  labs( x = "Average AUC", y = "Count")
#ggsave("supp.fig.5.png", dpi = 300, width = 175, height = 108, units = "mm")

@@

```{r, include=FALSE}
ggplot(cos.sim.auc.data, aes(x = cosine_similarity, y = average_auc))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs( x = "Cosine similarity between baseline COSMIC and FFPE signatures", y = "Average AUC")+
  geom_point(data = cos.sim.auc.outliers, color = "red")+
  geom_point(data = cos.sim.auc.outliers2, color = "purple")+
  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+
  geom_hline(yintercept=0.9, linetype="dashed", color = "purple")+
  geom_label_repel(data = cos.sim.auc.outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
  geom_label_repel(data = cos.sim.auc.outliers2, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
  geom_label_repel(data = cos.sim.sbs4, aes(label = signature), max.overlaps = 10, label.size =   0.1)

#ggsave("Fig2b.png", dpi = 300, width = 175, height = 108, units = "mm")

@@

```{r, include=FALSE}
set.seed(500)
#outer for loop: loop through each COSMIC v3 signature
average.sensitivity.values <- c()
average.specificity.values <- c()
for (x in 1:length(signature.names)){
  signature.sensitivity.vec <- c()
  signature.specificity.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
    classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df)
    sig.name <- class.df.colnames[3]
    #Count number of true positives (for sensititivity)
    sensitivity.df <- classification.df %>%
      filter(truth == sig.name & classify == sig.name)
    #Calculate sensitivity
    sensitivity.val = nrow(sensitivity.df)/100
    #Add to the vector of 10 sensitivity values
    signature.sensitivity.vec[y] = sensitivity.val
    #Count number of true negatives (for specificity)
    specificity.df <- classification.df %>%
      filter(truth == "FFPE" & classify == "FFPE")
    #Calculate specificity
    specificity.val = nrow(specificity.df)/100
    #Add to the vector of 10 specificity values
    signature.specificity.vec[y] = specificity.val
  }
  average.signature.sensitivity <- mean(signature.sensitivity.vec)
  average.sensitivity.values[x] <- average.signature.sensitivity
  average.signature.specificity <- mean(signature.specificity.vec)
  average.specificity.values[x] <- average.signature.specificity
}
```

```{r, include=FALSE}
sens.spec.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.sensitivity.values, average.specificity.values)

@@

ggplot(sens.spec.data, aes(x=sensitivity)) +
  geom_histogram(color="black", fill="lightblue")+
  labs( x = "Average sensitivity", y = "Count")
#ggsave("supp.fig.3.png", dpi = 300, width = 175, height = 108, units = "mm")

@@

ggplot(sens.spec.data, aes(x = cosine_similarity, y = sensitivity))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs( x = "Cosine similarity between baseline COSMIC and FFPE signatures", y = "Average sensitivity")+
  geom_point(data = outliers, color = "red")+
  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+
  geom_label_repel(data = outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)

#ggsave("supp.fig6.png", dpi = 300, width = 175, height = 108, units = "mm")

@@

ggplot(sens.spec.data, aes(x=specificity)) +
  geom_histogram(color="black", fill="lightblue")+
  labs( x = "Average specificity", y = "Count")
#ggsave("supp.fig.4.png", dpi = 300, width = 175, height = 108, units = "mm")

@@

ggplot(sens.spec.data, aes(x = cosine_similarity, y = specificity))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs( x = "Cosine similarity between baseline COSMIC and FFPE signatures", y = "Average specificity")+
  geom_point(data = outliers, color = "red")+
  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+
  geom_label_repel(data = outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)

#ggsave("supp.fig7.png", dpi = 300, width = 175, height = 108, units = "mm")

@@

```{r}
num.mut = 1000
step = seq(10, 990, by = 10)
```

@@

```{r}
sens.MMR.vec = evaluate_perf(100, length(step), test.MMR, "sens")
sens.tob.vec = evaluate_perf(100, length(step), test.tob, "sens")
```

@@

MMR.tob.sens <- ggplot(data = sens.df, mapping = aes(x = dilution, 
                                     y = sensitivity, 
                                     linetype = signature.sens)) +
  geom_point(aes(shape = signature.sens)) +
  geom_smooth(se=FALSE) +
  labs(x = "Percent of FFPE signature",
       y = "Sensitivity",
       linetype = "Baseline signature",
       shape = "Baseline signature") +
  coord_cartesian(xlim = c(0,100), ylim = c(0,1))+
  scale_y_continuous(breaks=seq(from=0, to=1, by=.2))+
  geom_hline(yintercept=0.8, linetype=4, color = "red")

#ggsave("fig3a.png", plot = MMR.tob.sens, dpi = 300, width = 175, height = 108, units = "mm")

@@

### Sensitivity threshold for all signatures

```{r}
#set threshold for a decent sensitivity for our classifier
optimal.sens = 0.8
max.sens.noise = c()
#loop through each signature
for (sig in 1:length(signature.matrix.list)){
  #empty lists to hold classifier outputs and ground truths
  probabilities.sig <- vector(mode = "list", length = length(step))
  truth.sig <- vector(mode = "list", length = length(step))

  #dilute each signature with some percent of FFPE
  for (x in 1:length(step)){
    sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x])
    ffpe.truth <- rep("FFPE", step[x])
    sample.sig <- create_signature_sample_vector(signature.matrix.list[[sig]],
                                                 num.mut-step[x])
    sig.truth <- rep("Biological", num.mut-step[x])
    #run classifier
    test <- get_classification_df(list(sample.sig, sample.ffpe), c("Biological","FFPE"),
                                  list(signature.matrix.list[[sig]], ffpe.sig))
    classification <- test %>%
      mutate(trueBinary = ifelse(truth == "Biological",1,0))
    #extract classifications and ground truth
    probabilities.sig[[x]] <- classification$Biological
    truth.sig[[x]] <- classification$trueBinary
  }

  #create ROC prediction object and evaluate
  test.sig <- prediction(probabilities.sig, truth.sig)
  sens.sig.vec = evaluate_perf(100, length(step), test.sig, "sens")

  #cut off noise past our threshold
  if(identical(which(sens.sig.vec <= optimal.sens), integer(0))){
    max.sens.noise[sig] = 100
    }
  else{
    max.sens.noise[sig] = min(which(sens.sig.vec <= optimal.sens))
    }
}
```

```{r}
#visualize results

all.sens.df <- data.frame(signature.names, max.sens.noise, ffpe.cosmic.cos.sims)
cos.sim.sbs4.sbs6 <- all.sens.df %>%
  filter(signature.names %in% c("SBS4", "SBS6"))
cos.sim.sens.outliers <- all.sens.df %>%
  filter(max.sens.noise <= 50)

all.sens <- ggplot(data = all.sens.df, mapping = aes(x = ffpe.cosmic.cos.sims,
                                         y = max.sens.noise))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  labs(x = "Cosine Similarity to FFPE",
       y = "Maximum Percent of FFPE Noise") +
  coord_cartesian(ylim = c(0,100), xlim = c(0,1)) +
  geom_hline(yintercept=50, linetype="dashed", color = "red")+
  geom_label_repel(data = cos.sim.sbs4.sbs6, aes(label = signature.names), 
                   max.overlaps = 10,
                   label.size =   0.1)+
  geom_label_repel(data = cos.sim.sens.outliers, aes(label = signature.names), 
                   max.overlaps = 10,
                   label.size =   0.1)

#ggsave("fig3c_sens.png", plot = all.sens, dpi = 300, width = 175, height = 108, units = "mm")
all.sens
```

```{r}
#linear relationship between cosine similarity and max FFPE noise
model.all.sens <- lm(max.sens.noise ~ ffpe.cosmic.cos.sims, data = all.sens.df)
summary(model.all.sens)
```

@@

```{r}
specificity = append(spec.MMR.vec, spec.tob.vec)
signature.spec = append(rep("SBS6", length(spec.MMR.vec)),
                        rep("SBS4", length(spec.tob.vec)))
dilution = append(step/10, step/10)
spec.df <- data.frame(dilution, specificity, signature.spec)

MMR.tob.spec <- ggplot(data = spec.df, mapping = aes(x = dilution, 
                                     y = specificity, 
                                     lty = signature.spec)) +
  geom_point(aes(shape = signature.spec)) +
  geom_smooth() +
  labs(x = "Percent of FFPE signature",
       y = "Specificity",
       lty = "Baseline signature",
       shape = "Baseline signature") +
  coord_cartesian(xlim = c(0,100), ylim = c(0,1))+
    scale_y_continuous(breaks=seq(from=0, to=1, by=.2))+
  geom_hline(yintercept=0.8, linetype=4, color = "red")

#ggsave("fig3b.png", plot = MMR.tob.spec, dpi = 300, width = 175, height = 108, units = "mm")

@@

### Specificity threshold for all signatures

```{r}
#set threshold for a decent specificity for our classifier
optimal.spec = 0.8
min.spec.noise = c()
#loop through each signature
for (sig in 1:length(signature.matrix.list)){
  #empty lists to hold classifier outputs and ground truths
  probabilities.sig <- vector(mode = "list", length = length(step))
  truth.sig <- vector(mode = "list", length = length(step))

  #dilute each signature with some percent of FFPE
  for (x in 1:length(step)){
    sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x])
    ffpe.truth <- rep("FFPE", step[x])
    sample.sig <- create_signature_sample_vector(signature.matrix.list[[sig]],
                                                 num.mut-step[x])
    sig.truth <- rep("Biological", num.mut-step[x])
    #run classifier
    test <- get_classification_df(list(sample.sig, sample.ffpe), c("Biological","FFPE"),
                                  list(signature.matrix.list[[sig]], ffpe.sig))
    classification <- test %>%
      mutate(trueBinary = ifelse(truth == "Biological",1,0))
    #extract classifications and ground truth
    probabilities.sig[[x]] <- classification$Biological
    truth.sig[[x]] <- classification$trueBinary
  }

  #create ROC prediction object and evaluate
  test.sig <- prediction(probabilities.sig, truth.sig)
  spec.sig.vec = evaluate_perf(100, length(step), test.sig, "spec")

  #cut off noise past our threshold
  if(identical(which(spec.sig.vec <= optimal.spec), integer(0))){
    min.spec.noise[sig] = 1
    }
  else{
    min.spec.noise[sig] = max(which(spec.sig.vec <= optimal.spec))
    }
}
```

```{r}
#visualize results

all.spec.df <- data.frame(signature.names, min.spec.noise, ffpe.cosmic.cos.sims)
cos.sim.sbs4.sbs6 <- all.spec.df %>%
  filter(signature.names == "SBS4")
cos.sim.spec.outliers <- all.spec.df %>%
  filter(min.spec.noise >= 50)

all.spec <- ggplot(data = all.spec.df, mapping = aes(x = ffpe.cosmic.cos.sims,
                                         y = min.spec.noise))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  labs(x = "Cosine Similarity to FFPE",
       y = "Minimum Percent of FFPE Noise") +
  coord_cartesian(ylim = c(0,100), xlim = c(0,1)) +
  geom_hline(yintercept=50, linetype="dashed", color = "red")+
  geom_label_repel(data = cos.sim.sbs4.sbs6, aes(label = signature.names), 
                   max.overlaps = 10,
                   label.size =   0.1)+
  geom_label_repel(data = cos.sim.spec.outliers, aes(label = signature.names), 
                   max.overlaps = 10,
                   label.size =   0.1)

#ggsave("fig3d_spec.png", plot = all.spec, dpi = 300, width = 175, height = 108, units = "mm")
all.spec
```

```{r}
model.all.spec <- lm(min.spec.noise ~ ffpe.cosmic.cos.sims, data = all.spec.df)
summary(model.all.spec)
```

@@

```{r, include=FALSE}
dilution.sig = signature.6.mat
dilution.sig.name = "SBS6"
total.mut = 2000
dilute = c(0.1,0.3,0.5,0.7,0.9)*total.mut
dilution.list <- vector(mode = "list", length = 5)
i = 1

@@

#ggsave("sig6reconstruction_1.png", plot = dilution.list[[1]], dpi = 300, width = 175, height = 243, units = "mm")
#ggsave("sig6reconstruction_2.png", plot = dilution.list[[2]], dpi = 300, width = 175, height = 243, units = "mm")
#ggsave("sig6reconstruction_3.png", plot = dilution.list[[3]], dpi = 300, width = 175, height = 243, units = "mm")
#ggsave("sig6reconstruction_4.png", plot = dilution.list[[4]], dpi = 300, width = 175, height = 243, units = "mm")
#ggsave("sig6reconstruction_5.png", plot = dilution.list[[5]], dpi = 300, width = 175, height = 243, units = "mm")
```

@@

```{r, include=FALSE}
#columns in data frame
sens = vector(mode = "list", length = length(signature.names))
spec = vector(mode = "list", length = length(signature.names))
acc = vector(mode = "list", length = length(signature.names))
signature = vector(mode = "list", length = length(signature.names))
